---
title: "Chapter 3's analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library("dplyr")
library("glue")
library("ggpubr")
library("lmerTest")
```

# Q2a (Section 3.5.1)

Load the dataframe for Q2a into the environment. The data comes from other processes/calculations and is already combined into the provided `.csv` files.
```{r}
naturalness = read.csv('naturalness.csv')
accuracy = read.csv('accuracy.csv')

# preprocess the data a bit
naturalness$type = factor(naturalness$type, levels = c("separate", "mapped"))
accuracy$type = factor(accuracy$type, levels = c("separate", "mapped"))
naturalness$sent_id = as.factor(naturalness$sent_id)
accuracy$sent_id = as.factor(accuracy$sent_id)
naturalness$same_big_family = as.factor(naturalness$same_big_family)
accuracy$same_big_family = as.factor(accuracy$same_big_family)
```

The following block corresponds to the results in Table 3.3

```{r, results="hold"}
languages = list('nl', 'fi', 'fr', 'es', 'ja')

for (lang in languages) {
    print(glue("LANGUAGE: {lang}"))
  
    # naturalness
    print("NATURALNESS")
    
    map = naturalness[(naturalness$source_lang == lang) & (naturalness$type == 'mapped'), ]$score
    sep = naturalness[(naturalness$source_lang == lang) & (naturalness$type == 'separate'), ]$score
    diff = median(map) - median(sep)
    value = wilcox.test(map, sep, paired=TRUE)$p.value
    
    # print out
    print(glue("Difference in median score (map - sep): {diff}, Wilcoxon test: {value}"))
    
    # accuracy
    print("ACCURACY")
    
    map = accuracy[(accuracy$source_lang == lang) & (accuracy$type == 'mapped'), ]$score
    sep = accuracy[(accuracy$source_lang == lang) & (accuracy$type == 'separate'), ]$score
    diff = median(map) - median(sep)
    value = wilcox.test(map, sep, paired=TRUE)$p.value
    
    # print out
    print(glue("Difference in median score (map - sep): {diff}, Wilcoxon test: {value}"))
}
```

The following block corresponds to the analysis that follows Table 3.3
```{r, results="hold"}

print("NATURALNESS")
model = lmer(score ~ type + (1|participant) + (1|sent_id), data=naturalness, REML=FALSE)
model.null = lmer(score ~ 1 + (1|participant) + (1|sent_id), data=naturalness, REML=FALSE)
anova(model, model.null)[2, 'Pr(>Chisq)']
fixef(model)['typemapped']

print("ACCURACY")
model = lmer(score ~ type + (1|participant) + (1|sent_id), data=accuracy, REML=FALSE)
model.null = lmer(score ~ 1 + (1|participant) + (1|sent_id), data=accuracy, REML=FALSE)
anova(model, model.null)[2, 'Pr(>Chisq)']
fixef(model)['typemapped']
```

# Q2b (Section 3.5.2)

The following blocks correspond to the analysis that precedes Table 3.4
```{r, results="hold"}

print("NATURALNESS")

model.null = lmer(score ~ 1 + (1|participant)  + (1|sent_id), data=naturalness, REML=FALSE)
model.1 = lmer(score ~ sent_similarity + (1|participant) + (1|sent_id), data=naturalness, REML=FALSE)
print(glue("ASPF ALONE, EFFECT: {fixef(model.1)['sent_similarity']}, SIGNIFICANCE: {anova(model.null, model.1)[2, 'Pr(>Chisq)']}"))
model.2 = lmer(score ~ same_big_family + (1|participant) + (1|sent_id), data=naturalness, REML=FALSE)
print(glue("SAME_FAMILY ALONE, EFFECT: {fixef(model.2)['same_big_family_True']}, SIGNIFICANCE: {anova(model.null, model.2)[2, 'Pr(>Chisq)']}"))
model.3 = lmer(score ~ same_big_family + sent_similarity + (1|participant) + (1|sent_id), data=naturalness, REML=FALSE)
print(glue("ASPF ALONE vs. BOTH, SIGNIFICANCE: {anova(model.1, model.3)[2, 'Pr(>Chisq)']}"))
print(glue("SAME_FAMILY ALONE vs. BOTH, SIGNIFICANCE: {anova(model.2, model.3)[2, 'Pr(>Chisq)']}"))

print("CONCLUSION")
print(glue("ASPF ALONE, EFFECT: {fixef(model.3)['sent_similarity']}"))
```

```{r, results="hold"}
print("ACCURACY")

model.null = lmer(score ~ 1 + (1|participant)  + (1|sent_id), data=accuracy, REML=FALSE)
model.1 = lmer(score ~ sent_similarity + (1|participant) + (1|sent_id), data=accuracy, REML=FALSE)
print(glue("ASPF ALONE, EFFECT: {fixef(model.1)['sent_similarity']}, SIGNIFICANCE: {anova(model.null, model.1)[2, 'Pr(>Chisq)']}"))
model.2 = lmer(score ~ same_big_family + (1|participant) + (1|sent_id), data=accuracy, REML=FALSE)
print(glue("SAME_FAMILY ALONE, EFFECT: {fixef(model.2)['same_big_family_True']}, SIGNIFICANCE: {anova(model.null, model.2)[2, 'Pr(>Chisq)']}"))
model.3 = lmer(score ~ same_big_family + sent_similarity + (1|participant) + (1|sent_id), data=accuracy, REML=FALSE)
print(glue("ASPF ALONE vs. BOTH, SIGNIFICANCE: {anova(model.1, model.3)[2, 'Pr(>Chisq)']}"))
print(glue("SAME_FAMILY ALONE vs. BOTH, SIGNIFICANCE: {anova(model.2, model.3)[2, 'Pr(>Chisq)']}"))

print("CONCLUSION")
print(glue("ASPF ALONE, EFFECT: {fixef(model.3)['sent_similarity']}"))
```

The following block corresponds to Table 3.4
```{r, results="hold"}
dataset_aspf = c(median(naturalness[(naturalness$source_lang == 'nl'), ]$phone_sim),
                median(naturalness[(naturalness$source_lang == 'fi'), ]$phone_sim),
                median(naturalness[(naturalness$source_lang == 'fr'), ]$phone_sim),
                median(naturalness[(naturalness$source_lang == 'es'), ]$phone_sim),
                median(naturalness[(naturalness$source_lang == 'ja'), ]$phone_sim))

print("NATURALNESS")

median_naturalness = c(median(naturalness[(naturalness$source_lang == 'nl'), ]$score),
                      median(naturalness[(naturalness$source_lang == 'fi'), ]$score),
                      median(naturalness[(naturalness$source_lang == 'fr'), ]$score),
                      median(naturalness[(naturalness$source_lang == 'es'), ]$score),
                      median(naturalness[(naturalness$source_lang == 'ja'), ]$score))

cor.test(median_naturalness, dataset_aspf, method='spearman')

print("ACCURACY")

median_accuracy = c(median(accuracy[(accuracy$source_lang == 'nl'), ]$score),
                      median(accuracy[(accuracy$source_lang == 'fi'), ]$score),
                      median(accuracy[(accuracy$source_lang == 'fr'), ]$score),
                      median(accuracy[(accuracy$source_lang == 'es'), ]$score),
                      median(accuracy[(accuracy$source_lang == 'ja'), ]$score))

cor.test(median_accuracy, dataset_aspf, method='spearman')
```